### 2021-04-24 assertions/doc/thesis-bernhardt/chapters/5-offline_design_tezos.tex
revision monolithic

\subsection{Monolithic orchestration}
In the monolithic approach, the assertion an parent code are merged into a single Smart Contract. This tightly couples and unites the code in one self-contained unit. The assertion code can be merged into the parent contract in two different ways. As shown in \figref{fig:mono_eps}, the assertions can be appended to the original contract as separate entrypoints. Alternatively, the assertion code can by prepended to the code of the respective entrypoint and thus become an integral part of the program (as shown in \figref{fig:mono_insert}).
\begin{figure}[!htb]
\subfloat[Appending assertions as separate entrypoints\label{fig:mono_eps}]{\includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep_basic.jpg}}
\quad
\subfloat[Inserting the assertion code into the original entrypoint\label{fig:mono_insert}]{\includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_basic.jpg}}
\caption{Approaches for the assembly of a monolithic contract}
\label{fig:monolithic_orchestration_basic}
\end{figure}

In variation (b), the link between the parent code and the respective assertion code is already established --- the assertion is executed automatically when invoking the entrypoint. Variation (a), on the other hand, still has to establish a link between the parent and the assertion entrypoint. It should be apparent, which of the entrypoints contains the original, and which one its respective assertion code. This can be solved by using fixed tag patterns for assertion entrypoints, which can be derived from the original entrypoint tag. \figref{fig:mono_eps} proposes the tag suffix ``\texttt{\_a}''. When an external transaction invokes such a monolithic contract, the bakers thus execute the contract normally. After completion, the baker then requests the validators to invoke the assertion separately. External transactions, of course, should only invoke the original entrypoints.

Both variations, however, do not yet provide any means to trigger several test runs per validator. Since only the contract itself knows how to calculate the required number of test runs $t$ (as per the range(s) of the random generator(s)), it cannot be passed to the validators as a parameter to the validation request. Instead, the contract itself needs to contain a controller to handle this. Similar to the assertion code, such a controller can be integrated into the parent entrypoint, or implemented as another separate entrypoint, as shown in \figref{fig:monolithic_orchestration}. Instead of invoking the assertion entrypoint (\texttt{A\_a}), the validators would then invoke the manager entrypoint (\texttt{A\_c}). This entryoint then determines $t$ and invokes the assertion entrypoint (\texttt{A\_a}) $t$ times as an internal operation. 
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep.jpg}
	\caption{Monolithic assembly with separate entrypoints for manager and assertion}
	\label{fig:monolithic_orchestration}
\end{figure}

@@

If a developer did not specify an assertion for all entrypoints of the parent contract, invoking their non-existing controller entrypoints would cause a transaction failure. To avoid this and thus increase the robustness, the compiler should generate dummy controllers for empty assertions. If the monolithic contract is assembled according to \figref{fig:mono_eps}, the dummy controller simply returns unit (i.e, an empty list of internal operations and the unchanged storage). For the variation given in \figref{fig:mono_insert}, the controller invokes the original entrypoint exactly once.

@@

For the variation of using separate entrypoints for the controllers and assertions, \figref{fig:interaction_monolithic} depicts the interactions of the nodes, validators and Smart Contracts on the blockchain. The execution model for distributed assertion checking requires a dedicated execution modus. A possible mechanism to activate this modus is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been validated. After the injection of a \texttt{txa(1, \%A, parameter)} operation, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, hence no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction to the network. In case of an assertion failure, the respective validator broadcasts the counterexample with the generated value to the network to be included in an upcoming block. The counterexample is thus validated like a normal transaction. As a consequence of the broadcast the counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be rejected. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

@@

\subsubsection{Evaluation}
The presented implementations of the monolithic approach each have strengths and weaknesses. Generally, they all share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly result in deeper parameter and code nestings, the code becomes less readable overall.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}

Another shortcoming, which applies to the implementation using separate entrypoints for controller and assertion, is the missing support of implicit entrypoint invocations. Even though the previously mentioned safety measures can prevent transactions failures, this still poses a restriction in the interaction with Smart Contracts.

@@

Besides these disadvantages, the monolithic approach has some distinct advantages. If controllers and assertions are implemented as separate entrypoints, they can be invoked independently --- this means, that the validators do not have to run the production code while checking the assertion. Conversely, the assertion code is not rerun if the parameter is valid and the transaction is broadcast to the network. This reduces the transaction cost for the user, as well as the overall footprint of the transaction on the network. If the assertion (and possibly controller) code is integrated into the parent entrypoint, this separation is not possible; however, the inseparableness the code in return avoids any bypassing of the assertion.

>>>>

\subsection{Monolithic orchestration}
In the monolithic approach, the assertion an parent code are merged into a single Smart Contract. This tightly couples and unites the code in one self-contained unit. As shown in \figref{fig:mono_eps}, the assertions are appended to the original contract as separate entrypoints. Since the \texttt{random} instruction is only available in the execution mode for assertion checking, the assertion code cannot be directly prepended to the respective production code. Otherwise, transactions invoking such an entrypoint could not be included into blocks, as these are validated in the normal execution mode.
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep_basic.jpg}
	\caption{Monolithic assembly - appending assertions as separate entrypoints}
	\label{fig:mono_eps}
\end{figure}

The interface of the contract should make apparent, which of the entrypoints contain production, and which contain their respective assertion code. This can be solved by using fixed tag patterns for assertion entrypoints, which can be derived from the production entrypoint tag. \figref{fig:mono_eps} proposes the tag suffix ``\texttt{\_a}''. When an external transaction invokes such a monolithic contract, the bakers thus execute the contract normally. After completion, the baker then requests the validators to invoke the assertion separately. External transactions, of course, should only invoke the production entrypoints.

The current version of the monolithic contract, however, do not yet provide any means to trigger several test runs per validator. Since the required number of test runs $t$ can only be determined at runtime, it cannot be passed to the validators as a parameter to the validation request. Instead, the contract itself should contain a controller to handle this. Similar to the assertion code, such a controller can be appended as another separate entrypoint, as shown in \figref{fig:monolithic_orchestration}. Instead of invoking the assertion entrypoint (\texttt{A\_a}), the validators would then invoke the controller entrypoint (\texttt{A\_c}), which determines $t$ and invokes (\texttt{A\_a}) $t$ times as an internal operation. 
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep}
	\caption{Monolithic assembly with separate entrypoints for cotroller and assertion}
	\label{fig:monolithic_orchestration}
\end{figure}

@@

If a developer did not specify an assertion for all entrypoints of the parent contract, invoking their non-existing controller entrypoints would cause a transaction failure. In order to avoid this and increase the robustness, the compiler should generate dummy controllers for empty assertions, which simply return unit (i.e, an empty list of internal operations and the unchanged storage).

@@

\figref{fig:interaction_monolithic} depicts the resulting interactions of the nodes, validators and monolithic Smart Contracts on the blockchain. The execution model for distributed assertion checking requires a dedicated execution modus. A possible mechanism to activate this modus is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been validated. After the injection of a \texttt{txa(1, \%A, parameter)} operation, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, hence no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction to the network. In case of an assertion failure, the respective validator broadcasts the counterexample with the generated value to the network to be included in an upcoming block. The counterexample is thus validated like a normal transaction. As a consequence of the broadcast the counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be rejected. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

@@

\subsubsection{Evaluation}
The presented implementations of the monolithic approach each have strengths and weaknesses. Generally, they share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly result in deeper parameter and code nestings, the code becomes less readable overall.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}

Another shortcoming is the missing support of implicit entrypoint invocations. Even though the previously mentioned measures can prevent transactions failures, this still poses a restriction in the interaction with Smart Contracts.

@@

Besides these disadvantages, the monolithic approach has some distinct advantages. The production and controller entrypoints they can be invoked independently, which means that the validators do not have to run the production code while checking the assertion. Conversely, the assertion code is not rerun if the parameter is valid and the transaction is broadcast to the network. This reduces the transaction cost for the user, as well as the overall footprint of the transaction in the network.
